---
title: "Fill in the blanks in `checkr`"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(checkr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.checker = checkr::check_for_learnr)
source(system.file("learnr_examples/internal-examples.R", 
                          package = "checkr"))
```

## Problem 1

> Fill in the blanks in the following code to create a `ggplot2` command that will produce the following scatter plot with the `mtcars` data.    
    ```{r echo = FALSE}
library(ggplot2)
ggplot(mtcars, aes(x = mpg, y = hp, color = cyl)) +
  geom_point()
```
>
> There are four blanks. You'll have to replace all of them with the correct contents to generate the plot.

```{r problem1, exercise = TRUE}
library(ggplot2)
ggplot(mtcars, aes(x = ____, y = ____, color = ____)) +
  ____()
```



To check the submission, we need to create a pattern that will let us look up the student's values for each of the blanks, then compare these to the correct answer. The `check_blanks()` function takes, as its second argument, an expression with the blanks. Since there are multiple blanks, each has been given a name for later reference. 
```{r problem1-check}
check_exer_14(USER_CODE)
```

## Problem 2

> Calculate the hypothenuse length C of a right triangle whose other edges are A and B.    

```{r pythag, exercise = TRUE}
A <- 3
B <- 4
C <- ........(A^2 + B^2)
```


```{r pythag-check}
check_pythag(USER_CODE)
```

Try some different submissions. 

- `C <- sqrt(A^2 + B^2)`
- `C <- log(A^2 + B^2)`
- `C <- ........(A^2 + B^2)` that is, the dots were not replaced
- `C <- sqrt(A + B)`


