---
title: "Introducing `checkr`"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(checkr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.checker = checkr::check_for_learnr)
# tutorial_options(exercise.checker = function(...) cat("Bogus checker\n"))

hello_check <- function(USER_CODE) {
  code <- for_checkr(USER_CODE)
  res <- misconception(code, line_where(code, is.null(F)), 
                       message = paste(
                         "Just typing the string doesn't do it. Note that the output", 
                         "has quotes and is formatted as an output, not a simple message."))
  res <- misconception(res, line_where(res, F == print), 
                       message = paste(
                         "Using print() isn't right.",
                         "True, that's the way character strings are displayed,", 
                         "but the format of the display is as a quoted output and",
                         "not just a simple message."))
  res <- line_where(res, F == cat, message = "Try the cat() function.")
  check(arg_number(res, 1), passif(V == "Hello, World!", "Good job!"),
        failif(tolower(V) == "hello, world!", "Get the capitalization right!"),
        failif(TRUE, "The string should be 'Hello, World!', not '{{V}}'.")) 
}

two_plus_two_check <- function(USER_CODE) {
  res <- for_checkr(USER_CODE)
  res <- line_where(res, V == 4, message = "The result should be 4, not {{V}}.")
  res <- line_where(res, F == `+`, 
                    message = 
                      paste("Think about what function corresponds to 'addition'.", 
                            "It isn't {{F}}."))
  arg_number(res, 1, failif(V != 2, "The first argument should be 2, not {{EX}}."),
             passif(TRUE, "Yes, that's it!"))
}

hello_fun_strict <- function(USER_CODE) {
  code <- for_checkr(USER_CODE)
  line_binding(code, 
               cat("Hello, World!"),  # a pattern with no flexibility.
               passif(TRUE, "That's right."),
               fail = "No. Try aain.")
}
```

## What's this document about?

This is a `learnr` tutorial to display some examples of code-checking with the `checkr` package.

As you may know, `learnr` is a system for creating and distributing interactive documents, particularly documents that provide exercise-boxes where the reader can type and execute R code.

`Checkr2` is a package that provides feedback about the code that is being executed in a `learnr` exercise box.

## Hello, World!

"Hello, World!" is a famous introductory example in programming. In this task, you will ... 

> Write the one-line version of "Hello, World!" in R. Your code should cause the message `Hello, World!` to appear. 

Press "Run" to try out your command to make sure it does what you were expecting. When you are satisfied, press "Submit" to get the `checkr` feedback.

The box below has been instrumented with `checkr` statements intended to guide the user to a correct answer. Exactly what "guide the user" means is ambiguous. Play around with various possible answers and see if you agree that the `checkr` statement I choose are helpful.


```{r eval = FALSE, echo = FALSE}
hello_check(quote("\"Hello, World!\""))
hello_check(quote(cat("Hello, World!")))
hello_check(quote(print("Hello, World")))
hello_check(quote(cat("hello")))
hello_check(quote(cat("Hello, world!")))
```

```{r hello-world-check}
hello_check(USER_CODE)
```

```{r hello-world, exercise = TRUE}
"Hello, World!"
```

### The man behind the curtain

Ordinarily, you would not show students the `checkr` statements implementing this behavior. But our purpose here is to introduce `checkr`, So here are the statements for the above exercise.

```{r echo = FALSE, comment = "", tidy = FALSE}
print_function_contents(
  hello_check,
  from_file = system.file("learnr_examples/internal-examples.R", 
                          package = "checkr"))
```

Breaking this down, line by line:

- [1] accepts the user submission from `learnr`. This is always called `USER_CODE`. The function `for_check()` does some pre-processing of the user submission to turn it into evaluated code and format it for use in later `checkr` functions.
- [2] tests for a particular kind of mistaken answer. The `misconception()` function will generate a `checkr` fail message, if the pattern identified in the second argument passes. That pattern, `line_where(code, is.null(F))`, means, "scan the code looking for a line where no function is being used." This will captures a line that contains only a character string.
- [5] tests for another specific misconception, that the user invokes `print()` on the string. 
- [9] looks whether the function invoked by the user is `cat()`. If not, the check fails. (Note that [5] already ruled out that `print()` was being invoked.)
- [10] checks the argument to the `cat()` function. (We know it's `cat()`, because [9] has established this.) If that argument is exactly `"Hello, World!" the submission passes. Otherwise, we check for a particular error involving capitalization and, if that's not the case, generate a message to tell the student what's wrong.

Depending on the submission, any of the checks on lines 2, 5, 9, and 10 might fail. If a check fails, later checks that use the previous result will short circuit to a failed check. For instance, if the check on line [2] fails, the remaining checks won't be performed in detail: they will just pass along the failed result from line [2].

An instructor with a different pedagogical approach might prefer to structure the checking in an entirely different way. For instance, here are `checkr` statements that simply tell the user whether or not the submission did what was requested:

```{r echo = FALSE}
print_function_contents(
  hello_fun_strict,
  from_file = system.file("learnr_examples/internal-examples.R", 
                          package = "checkr"))
```

```{r hello-world-strict, exercise = TRUE}
"Hello, World!"
```

```{r hello-world-strict-check}
hello_fun_strict(USER_CODE)
```

## Two plus two

Not as famous as "Hello, World!" is the first example in the `learnr` documentation:

> Write the code to add two and two.


```{r two-plus-two, exercise=TRUE}

```

Of course, a submission of `2 + 2` will work. See what happens when you try to fool the checking system with a submission like `6 - 4` or `1 + 1 + 1 + 1`. 


```{r two-plus-two-code-check}
1 # To trigger pre-check, this chunk must have something in it.
```

```{r two-plus-two-check}
two_plus_two_check(USER_CODE)
```

Here are the `checkr` statements implementing the above behavior:
```{r echo = FALSE}
print_function_contents(
  two_plus_two_check,
  from_file = system.file("learnr_examples/internal-examples.R", 
                          package = "checkr"))
```


