---
title: "Matching patterns with `redpen`"
author: "Danny Kaplan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Matching patterns with redpen}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document explores possibilities for using the `redpen` package as the basis for code pattern matching.

## 2 + 2

> Write a command to add two and two.

Criteria:
- result should be 4
- the `+` operator should be used
- both arguments should be 2

```{r}
good1 <- quote(2 + 2)
good2 <- quote((1+1) + 2)
bad1 <- quote(3 + 1)
bad2 <- quote(5 - 1)
bad3 <- quote(2 * 2)
bad4 <- quote(2 - 2)
```

- Result is 4. 
```{r}
pattern <- ..(foo) ~ identical(foo, 4)
redpen::redpen::node_match(good1, !!pattern)
redpen::node_match(bad3, !!pattern)
redpen::node_match(bad4, !!pattern)
```
- `+` operator is used
```{r}
op <- `+`
pattern <- ..(foo)(...) ~ identical(foo, `+`)
redpen::node_match(good1, !!pattern)
```
Note that the `..` is used to get the value of `foo` rather than the (??tag??, ??name??, what?).

- Both arguments are two.
```{r}
pattern <- .(op)(.(aa), .(bb)) ~ aa == bb
pattern2 <- .(op)(..(aa), ..(bb)) ~ aa == bb # the values
redpen::node_match(good1, !!pattern)
redpen::node_match(good1, !!pattern)
redpen::node_match(good2, !!pattern)
redpen::node_match(good2, !!pattern2)
redpen::node_match(bad1, !!pattern)
redpen::node_match(bad2, !!pattern)
redpen::node_match(bad3, !!pattern)
redpen::node_match(bad4, !!pattern)
```

## Some possible supervisor functions


Idea: Use moustache interpolation of bindings.

```{r}
fail <- function(m = "Feedback message",  ...) {
  # immediately terminate if any of the ... are FALSE
  
}
note <- function(m = "Point out so and so.", ...) {
  # make a note of a potential problem
  # to be reported along with the failure message
  # if the fail() or default() functions are called.
  
  # maybe capture the failed component of ... and
  # include that in the message.
}
misconception <- function(m = "That's a common misconception.", ...) {
  # A kind of `fail()` to make explicit that this is 
  # a mistake anticipated by the author

}
pass <- function(m = "Good job!", ...) {
  # immediately report success if all of the ... are true
  # this proactive pass requires success for ...
  
}
default <- function(m = "Didn't see anything wrong") {
  # return a passing feedback, but include the notes.
}
roughly <- function(x, y, epsilon = 0.1) {
  abs(x - y) < epsilon
}
has_name <- function(nm, set) {
  if (is_lang(set)) set <- all.names(set)
  nm %in% set
}
```

## Degrees and radians

> Calculate the cosine of 81 degrees.

```{r}
good <- quote(cos(81 * pi / 180))
bad1 <- quote(cos(81))
bad2 <- quote(cos(81 / 180))
```

- Converted to radians correctly
```{r}
pattern1 <- cos(..(foo)) ~ roughly(foo, 81*pi/180)
redpen::node_match(good, !!pattern)
redpen::node_match(bad1, !!pattern)
redpen::node_match(bad2, !!pattern)
```
- Pi appears in the argument to cosine
```{r}
pattern2 <- cos(.(foo)) ~ has_name("pi", all.names(foo))
redpen::node_match(good, !!pattern2)
redpen::node_match(bad1, !!pattern2)
redpen::node_match(bad2, !!pattern2)
```
Note that `.(foo)` has only one dot, since we want the expression rather than the value.

- Can we figure out if pi is on the top or bottom?
```{r}
pattern <- .(a) / .(b)  ~ has_name("pi", a)
redpen::node_match(quote(81 * pi/180) , !!pattern)
```

## Assignment

Patterns to pull out the LHS and RHS of assignment.
```{r}
y <- 7
statement <- quote(x <- y^2)
```

```{r}
pattern_name <- `<-`(.(a), .(b)) ~ a
pattern1 <- `<-`(., .(b)) ~ b
pattern2 <- `<-`(.(a), ..(b)) ~ b
redpen::node_match(statement[[3]], !!pattern_name)
redpen::node_match(statement, !!pattern1)
redpen::node_match(statement, !!pattern2)
```

A try at an easy-to-use "assigns_to()" function. This generates a formula that can be used in `redpen::node_match()` and returns `TRUE` if the statement assigns to.

```{r}
assigns_to <- function(ex, assignable) {
  x <- rlang::node_cadr(rlang::enquo(assignable))
  pattern <- `<-`(lhs, ...) ~ TRUE
  mut_node_cadr(rlang::f_lhs(pattern), x) #in place mutation
  
  redpen::redpen::node_match(ex, !!pattern)
}
get_assigns_val <- function(ex, assignable) {
  x <- rlang::node_cadr(rlang::enquo(assignable))
  pattern <- `<-`(lhs, ..(a)) ~ a
  mut_node_cadr(rlang::f_lhs(pattern), x) #in place mutation
  
  redpen::redpen::node_match(ex, !!pattern)
}
get_assigns_expr <- function(ex, assignable) {
  x <- rlang::node_cadr(rlang::enquo(assignable))
  pattern <- `<-`(lhs, .(a)) ~ a
  mut_node_cadr(rlang::f_lhs(pattern), x) #in place mutation
  
  redpen::redpen::node_match(ex, !!pattern)
}
get_arg_expr <- function(arg_nm) {
  x <- rlang::node_cadr(rlang::enquo(arg_nm))
  pattern <- .(foo)(arg_nm = .(a), ...) ~ a
  
  first <- rlang::node_cadr(pattern)
  second <- rlang::node_cdr(first)
  names(second) <- as.character(x)
  return(pattern)
}
get_arg_val <- function(arg_nm) {
  x <- rlang::node_cadr(rlang::enquo(arg_nm))
  pattern <- .(foo)(arg_nm = ..(a), ...) ~ a
  
  first <- rlang::node_cadr(pattern)
  second <- rlang::node_cdr(first)
  names(second) <- as.character(x)
  return(pattern)
}
get_fun_name <- function() {
  .(fn)(...) ~ fn
}

# look over the common patterns and customize accordingly
# e.g. explicit data argument
get_data_value <- function(ex) {}
get_data_name <- function(ex) {}
get_formula <- function(ex){} 
```


## Chains

```{r}
library(dplyr)
ex <- quote(
  mtcars %>% 
  group_by(cyl) %>%
  mutate(ghp = mpg / hp) %>%
  summarise(mn = mean(ghp))
)
eval(ex)
ex2 <- quote(mtcars %>% summarise(m = mean(hp)))
```

```{r}
is_chain <- function(ex) {
  if (is.name(ex) ||
      is.pairlist(ex)) {
    FALSE
  } else {
    identical(as.name(rlang::node_car(ex)), as.name("%>%"))
  }
}

# convert a chain into a list of expressions
chain_elements <- function(ex) {
  if (is_chain(ex)) {
    c(chain_elements(rlang::node_cadr(ex)),
      chain_elements(rlang::node_cddr(ex)))
  } else {
    ex
  }
}

# Given a list of chain elements, turn them back into
# a chain

elements_to_chain <- function(elements) {
  if (length(elements) == 1) return(elements[[1]])
  chain_start <-
    rlang::lang(quote(`%>%`), 
         elements[[1]], 
         elements[[2]])
  for (el in elements[-(1:2)]) {
    chain_start <- rlang::lang(quote(`%>%`),
                        chain_start,
                        el)
  }
  chain_start
}

# evaluate an element of the chain at a given input
eval_chain_element <- function(input, ex) {
  with_input <- rlang::lang(quote(`%>%`),as.name("input"), ex)
  browser()
  eval(with_input)
}

```


```{r}
# pull out everything in the chain *after* the specified function or name
chain_head <- function(chn, pat, include.match = TRUE) {
  pat <- rlang::node_cadr(rlang::enquo(pat))
  pattern <- rlang::new_formula(lhs = pat, rhs = TRUE)
  if ( ! is_chain(chn)) stop("not a chain")
  parts <- chain_elements(chn)
  # Find a matching part of the chain
  kfound <- 0
  for (k in 2:length(parts)) {
    if ( ! is_empty(redpen::redpen::node_match(parts[[k]], 
                               !!pattern))) {
      kfound <- k
      break
    }
  }
   # Failed to find any
  if (kfound == 0) return(NULL)
  kfound <- ifelse(include.match, kfound, kfound-1)

  elements_to_chain(parts[1:kfound])
}
# give back a headless chain
chain_tail <- function(chn, pat, include.match = TRUE){
    pat <- rlang::node_cadr(rlang::enquo(pat))
  pattern <- rlang::new_formula(lhs = pat, rhs = TRUE)
  if ( ! is_chain(chn)) stop("not a chain")
  parts <- chain_elements(chn)
  # Find a matching part of the chain
  kfound <- 0
  for (k in 2:length(parts)) {
    if ( ! is_empty(redpen::redpen::node_match(parts[[k]], 
                               !!pattern))) {
      kfound <- k
      break
    }
  }
   # Failed to find any
  if (kfound == 0) return(NULL)
  kfound <- ifelse(include.match, kfound, kfound+1)

  elements_to_chain(parts[-(1:(kfound-1))])
}
  

# pull out the input data or output data for the 
# command set in the pattern
chain_value_before <- function(chn, pat){
  pat <- rlang::node_cadr(rlang::enquo(pat))
  pattern <- rlang::new_formula(lhs = pat, rhs = TRUE)
  if ( ! is_chain(chn)) stop("not a chain")
  parts <- chain_elements(chn)
  # Find a matching part of the chain
  kfound <- 0
  for (k in 2:length(parts)) {
    if ( ! is_empty(redpen::redpen::node_match(parts[[k]], 
                               !!pattern))) {
      kfound <- k
      break
    }
  }
 
  
  # Failed to find any
  if (kfound == 0) return(NULL)
  # Found one
  val <- eval(parts[[1]])
  for (k in 2:kfound) {
    val <- eval_chain_element(val, parts[[k]])
  }    
    
  val
}

chain_value_after <- function(chn){}

# pull out the expression immediately before or after
# the one that matches the pattern
chain_expr_before <- function(chn){}
chain_expr_after <- function(chn){}
```
