---
title: "Examples for `checkr`"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(checkr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.checker = checkr::check_for_learnr)
# tutorial_options(exercise.checker = function(...) cat("Bogus checker\n"))
```






## Using `rep()`

a. Create a vector named `Id` that has contents `1 1 1 2 2 2 3 3 3 4 4 4`. Use the `rep()` function to do this.
```{r bunch-1-a, exercise = TRUE, echo = FALSE}

```

```{r bunch-1-a-solution, echo = FALSE}
Id <- rep(1:4, each = 3)
```

```{r bunch-1-a-hint-1, echo = FALSE}
1:4
```

```{r bunch-1-a-hint-2, echo = FALSE}
rep(1:4, .... = 3) 
```

```{r bunch-1-a-code-check, echo = FALSE}
1 # There must be something in this chunk.
```

```{r bunch-1-a-check, echo = FALSE}
checkr:::rep_1245(USER_CODE)
```


b. Create a vector named `Letter` that contains `"a" "b" "c" "a" "b" "c" "a" "b" "c" "a" "b" "c"`
```{r bunch-1-b, exercise = TRUE, echo = FALSE}

```

```{r bunch-1-b-solution, echo = FALSE}
Letter <- rep(c("a","b","c"), each = 4)
```

```{r bunch-1-b-hint-1, echo = FALSE}
c("a", "b", "c")
```

```{r bunch-1-b-hint-2, echo = FALSE}
rep(...., length.out = 12)
```

```{r bunch-1-b-code-check, echo = FALSE}
1 # chunk can't be empty
```

```{r bunch-1-b-check, echo = FALSE}
checkr:::rep_abcd(USER_CODE)
```


c. You've constructed `Id` in the previous exercise. Copy the statement you used to do that into this code block. Then create two more vectors:
    1. `x` which will be evenly spaced numbers between 1 and 43. `x` will have the same length as `Id`.
    2. `y` which will be similar to `x`, but will consist of evenly spaced numbers between -20 and 0.
    
    Once you have constructed `x` and `y`, make a data frame called `df` that contains `x`, `y`, `Id` as columns. The variable names should be `x`, `y`, and `Id`.

    ```{r bunch-1-c, exercise = TRUE, echo = FALSE}
Id <- rep(1:4, each=3) 
x <- seq(1, 43, length = length(Id))
y <- seq(-20, 0, length = length(Id))
df <- data.frame(x=x, y = y, Id = Id)
```

```{r bunch-1-c-solution, echo = FALSE}
Id <- rep(1:4, each = 3)
x <- seq(1, 43, along.with=Id)
y <- seq(-20,0, along.with=Id)
df <- data.frame(x = x, y = y, Id = Id)
```

```{r bunch-1-c-check, echo = FALSE}
#USER_CODE <- "Id <- rep(1:4, each=3); x <- seq(1, 43, length = length(Id)); y <- seq(-20, 0, length = length(Id)); df <- data.frame(x=x, y = y, Id = Id)"
checkr:::df_abcd_1234_x_y(USER_CODE)
```

